Dynamics of Anthropomorphism in Human-Robot Interaction

### 2014-08-27 paper.tex
Full rewrite by Julia -- submitted version

%\documentclass{frontiersENG} % for Engineering articles
\documentclass{frontiersSCNS} % for Science articles
%\documentclass{frontiersMED} % for Medicine articles
%\documentclass{frontiersFPHY} % for Physics articles

%\usepackage[utf8x]{inputenc}
%\usepackage[T1]{fontenc}

%\usepackage[english]{babel}

\usepackage{graphicx}
\graphicspath{{figs/}}
\usepackage{tikz}
\usetikzlibrary{shapes}
\usetikzlibrary{svg.path}
%
\usepackage{amsmath} 
\usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
\usepackage{color, soul}
\usepackage{url}
\usepackage{multirow}
\usepackage{array}
\usepackage{fixltx2e}
\usepackage{textcomp}

\usepackage{caption}
\usepackage{subcaption}
\usepackage{float}
\floatstyle{plaintop}

\usepackage{paralist} % inparaenum

\usepackage[bookmarks]{hyperref}

\usepackage[draft, nomargin, marginclue, footnote]{fixme}
\fxsetup{targetlayout=color}

@@

\usepackage{xifthen}% provides \isempty test

% use either \Ant or \Ant[R,H,C,...]
\newcommand{\Ant}[1][]{%
      \ifthenelse{\isempty{#1}}%
        {$\mathcal{A}$}
        {$\mathcal{A}(#1)$}
}

\newcommand{\AntNorm}[1][]{%
      \ifthenelse{\isempty{#1}}%
      {$\widehat{\mathcal{A}}$}
      {$\widehat{\mathcal{A}}(#1)$}
}

% for use inside a math environment
\newcommand{\ant}[1][]{%
      \ifthenelse{\isempty{#1}}%
        {\mathcal{A}}
        {\mathcal{A}(#1)}
}

\newcommand{\antNorm}[1][]{%
      \ifthenelse{\isempty{#1}}%
      {\widehat{\mathcal{A}}}
      {\widehat{\mathcal{A}}(#1)}
}

\newcommand{\ICA}{{$\mathcal{A}_0$~}}
\newcommand{\SLA}{{$\mathcal{A}_\infty$~}}
\newcommand{\sla}{{\mathcal{A}_\infty}}
\newcommand{\AntMax}{{$\mathcal{A}_{max}$~}}
\newcommand{\antMax}{{\mathcal{A}_{max}}}

@@

\def\journal{Cognitive Science}%%% write here for which journal %%%
\def\DOI{}
\def\articleType{Research Article}
\def\keyFont{\fontsize{8}{11}\helveticabold }
\def\firstAuthorLast{Fink {et~al.}} %use et al only if is more than 1 author
\def\Authors{Julia Fink\,$^{1*}$, S\'{e}verin Lemaignan$^{1}$, Claire Braboszcz$^{2}$ and Pierre Dillenbourg$^{1}$}
\def\Address{$^{1}$Computer-Human Interaction in Learning and Instruction (CHILI) \\ Ecole Polytechnique F\'{e}d\'{e}rale de Lausanne (EPFL) \\ CH-1015 Lausanne, Switzerland \\
\vspace{0.25cm}
$^{2}$Laboratory for Neurology and Imaging of Cognition (LabNIC) \\
Universit\'{e} de Gen\`{e}ve \\ CH-1211 Gen\`{e}ve, Switzerland 
}
\def\corrAuthor{Julia Fink \vspace{1mm}}
\def\corrAddress{EPFL - CHILI, RLC D1 740, Station 20, CH-1015 Lausanne, Switzerland}
\def\corrEmail{julia.fink@epfl.ch}

@@

\textit{Anthropomorphism} -- the people's tendency to perceive human-like
characteristics in non-human agents -- is a phenomenon that is most often
studied from the sole perspective of the human-like \emph{design} of robots.

By building upon experimental results from several studies, a comprehensive
synthesis of the literature, as well as insights from social psychology,
cognitive science, and neuroscience, this article proposes to extend our current
understanding of anthropomorphism into a \textbf{formal model} that defines a
\textbf{clear terminology}, accounts for the \textbf{non-monotonic dynamics} of
the phenomenon, and relate the observed \emph{effects} of anthropomorphism to
their \textbf{underlying cognitive precursors}.

We also examine the \textbf{factors} that determine anthropomorphism (the
design of the non-human agent, the individual psychological determinants of the
human user and the context of use) and we discuss several \textbf{episodes of
interaction} that influence the anthropomorphic projections on the robot, like
the \textbf{novelty effect}, the role of \textbf{familiarization} and the
impact of \textbf{disruptive behaviors}.

The article also discusses techniques and scenarios to measure anthropomorphism,
and we conclude with a few suggestions on how others can apply our model to
better describe our tendency to perceive human-like characteristics in non-human
agents.

\tiny
 \keyFont{ \section{Keywords:} Anthropomorphism, Design, Human-Robot Interaction, Social Issues in Robotics, Acceptance of Robots}

\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
%
%		INTRODUCTION & OVERVIEW
%
%
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Introduction}
\label{sec:intro}

The Uncanny Valley hypothesis \citep{mori_uncanny_1970} proposes that human-like
designs of objects and non-human agents can evoke both positive and negative feelings in the human observer / user. A positive valence reflects the experience of 
emotional engagement with and feelings of empathy for the non-human agent
\citep{cheetham_human_2011}. This in turn may lead to an effective (human-like and 
human-social) interaction with the non-human agent, and increase people's acceptance of
it. 
%
%However, at a certain point as the non-human agent
%approaches but fails to closely resemble a human, the evoked valence in the
%human user suddenly turns into a negative uncanny feeling of what
%\cite{mori_uncanny_1970} describes as disgust and revulsion. This, in turn,
%might lead to rejection of the non-human agent. The Uncanny
%Valley hypothesis predicts that with further increasing the human-likeness of
%the non-human agent such that it appears so realistic that it is not
%distinguishable from a human but perceived to be human, the effect is again a
%positive one, reflecting feelings and engagement that are similar as those
%evoked by a human-human interaction. 
%Of course, developers and designers of artificial agents
%(\eg virtual characters, robots) try to exploit the positive effects of 
%\textbf{anthropomorphic form} and avoid ``falling'' into this Uncanny Valley.
%Consequently, there is a considerable research effort to
%better understand the boundaries of anthropomorphic design and their effects
%on the interaction and experience. To date, findings have been however
%inconsistent \citep{cheetham_human_2011}.

A phenomenon that is often studied related to this is \textbf{anthropomorphism}.
Along this line, this article aims at fostering a better understanding of
anthropomorphism (noted \Ant in this paper) in human-robot interaction 
by considering the phenomenon
as a whole. Robotics researchers often tend to believe that
anthropomorphism only describes a static set of human-like features of a robot
(like shape, speech capabilities, facial expressions, etc.). We refer to these
characteristics as the \emph{anthropomorphic design} of the
robot~\citep{fink_anthropomorphism_2012}, and propose that \emph{anthropomorphism} in 
general refers to a \emph{social phenomenon} that emerges from the (real or
imagined) \emph{interaction} between a non-human agent (\eg a robot) and a 
human~\citep{persson_anthropomorphism_2000}. According to~\cite{epley_when_2008},
anthropomorphism includes that the human perceives, for instance, emotional states, motivations, or 
intentions in the non-human agent, and tends to ascribe those qualities to it.
As such, the dynamic and
socio-cognitive dimensions of anthropomorphism are essential to understand the
complex bonds that human users build with artificial agents such as robots.

\subsection{A few results from the field}
\label{sec:field-results}

To support the reflection on variances and dimensions of anthropomorphism, it may be 
useful to consider a few experimental
results. We studied anthropomorphism in several short- and long-term
human-robot interaction studies that we carried out with embodied robots, both
in ecologically valid real-world environments and controlled lab settings. Those
experiments are presented in details in previously published
articles~\citep{fink_anthropomorphic_2012, fink_living_2013, fink2014which,
lemaignan2014dynamics}. We only highlight here a few qualitative results that
underline the complexity of the phenomenon.

\begin{figure}[b]
        \centering
        \begin{subfigure}[t]{0.48\columnwidth}
                \includegraphics[width=\textwidth]{roomba-activities.png}
                \caption{Activities related to Roomba, n = 9 households}
                \label{fig:roomba-activities}
        \end{subfigure}%
        \hspace{0.5cm} 
        \begin{subfigure}[t]{0.48\columnwidth}
                \includegraphics[width=\textwidth]{roomba-perception.pdf}
                \caption{People's perception of Roomba, n = 15 participants}
                \label{fig:roomba-perception}
        \end{subfigure}
    \caption{Usage and perception of the Roomba vacuum-cleaning robot over a 6 months period. Qualitative data suggest that people's activities related to the robot as well as their perception of the robot change over time.}
    \label{fig:roomba}
\end{figure}

Figures~\ref{fig:roomba-activities} and~\ref{fig:roomba-perception} give a first
perspective on how naive users interact and perceive robots over a long period
of time. Here, \emph{Roomba} vacuum cleaning robots were given to nine
households for a 6-months period. The aim was to explore how people use and
perceive the robot over time, and how they generally live with the Roomba in
their home. At various time points, users were asked to rate their perception of
different features of the robot on a 7-point rating scale. As
Figure~\ref{fig:roomba-perception} shows, some features remained stable over
time (\emph{ease of use} and \emph{fun}), and others decreased: the perceived
\emph{usefulness}, \emph{intelligence} and \emph{impact} on their household.
\emph{intelligence} and \emph{impact} reflect the cognitive engagement of the
user towards the machine: we can assume that one does not care to engage in a
(cognitive) interaction with an artifact perceived as non-intelligent and with
no impact on one's life (and the level of activity pictured in
Figure~\ref{fig:roomba-activities} qualitatively confirms this intuition). Even
if one can argue that the Roomba is not designed to foster interaction, it
raises a first question: how to sustain engagement between a robot and a human
over a long period of time? And, as a pre-requisite to answer, do we actually
understand the psychological phenomenon, and can we effectively account for the
dynamics of these interactions?

Those questions are certainly not new, but paradoxically, while anthropomorphism
is a commonly discussed trait of human-robot interaction, it appears
that we lack formal grounds and models that would account for the long-term
dynamics of these interactions.

\begin{figure}[b]
        \centering
        \begin{subfigure}[b]{0.48\columnwidth}
                \centering
                \includegraphics[width=0.75\textwidth]{ranger}
                \caption{Interaction scenario of the \emph{Ranger} domino study: the robot toy box Ranger transports domino tiles between the two children}
                \label{fig:ranger-expe}
        \end{subfigure}%
        \hspace{0.5cm}
        \begin{subfigure}[b]{0.48\columnwidth}
            \includegraphics[width=\textwidth]{ranger-interactions-vs-qualitative-score}
            \caption{We found a significant negative correlation ($p=0.048$) between the
    number of interactions and the qualitative anthropomorphic projections. n = 13 groups}
            \label{fig:qualitative-score}
        \end{subfigure}
        \caption{Children's interaction with Ranger, and their anthropomorphic projections onto the robot. Overall, large variations of both children's behavior and their perception of the robot were found but seemed to be correlated to each other.}
    \label{fig:ranger}
\end{figure}

%\fixme{It was written as caption of fig:ranger: ``Level of interaction with a semi-anthropomorphic robot is \emph{not} a predictor of anthropomorphic projections.'' Comment JF: this statement does not make sense to me as we found a significant negative correlation between the two.}

Figure~\ref{fig:qualitative-score} calls for other initial remarks. This diagram
shows data from a (yet unpublished) study in which 13 pairs of children (4-5 years old)
played for about 20 minutes a Domino game with a robot toy box (the robot was tasked to
carry domino pieces from one child to the other, Figure~\ref{fig:ranger-expe}).
The $x$-axis represents the number of social interactions between the children
and the robot (like talking to the robot or showing objects), based on video annotations. On
the $y$-axis, we have a synthetic qualitative measurement of anthropomorphic
projections onto the robot, based on two interviews with the children. The correlation between both these
values is significantly \emph{negative}: the more the children engaged into
interaction, the less prone to anthropomorphic projections they were. This
diagram also highlights an interesting cluster of children who did not interact
much with the robot, and still reported a strong anthropomorphic perception of
the robot.

These results may seem counter-intuitive at first sight, and illustrate that the
mechanisms that relate anthropomorphism to adoption and engagement are
non-trivial. In section~\ref{sec:cognition-neuroscience}, we sketch a model of
the cognitive correlates of anthropomorphism that partially account for these
results.

\subsection{Towards a formal understanding of anthropomorphism}

These first remarks lead to the main question we attempt to address in this
article: how can we understand \emph{anthropomorphism} as a whole, and in such a
way that it supports or improves human-robot interaction?

We propose to discuss anthropomorphism from three
complementary perspectives:

%Based on an extensive literature review~\citep{fink_anthropomorphism_2012},
%several field and lab studies previously mentioned, and insights from the
%cognitive neuroscience, we propose to discuss anthropomorphism from three
%complementary perspectives.

First, by reviewing what anthropomorphism actually \emph{means}:
besides an \emph{artifact-centered} understanding that explains anthropomorphism from the human-like design of the non-human artifact, it seems important to explicit the social component, namely a \emph{human-centered} understanding, of anthropomorphism.

Second, we propose a model, called the \textit{Dynamics of Anthropomorphism} to understand how anthropomorphism evolves over time ~\citep{lemaignan2014dynamics}.
These dynamics are non-monotonic, and effects like the
\emph{novelty effect} play here a key role.

Finally, we adopt a third perspective on anthropomorphism by investigating its
\emph{cognitive correlates}. Several studies previously showed that we build
instinctive bonds with robots as soon as we attribute animacy, and we also know
from social psychology some of the mechanisms that lead to
\emph{familiarization}. We attempt here to ``connect the dots'', in order to
build a practical model of the cognitive stages we encounter in human-robot
interaction.

%\paragraph{Two hypotheses}
Our research question also translates to two
hypotheses, which structure the reflections of this article:

\begin{enumerate} 
\item ~A person's tendency to anthropomorphize a robot
    is based on several factors, including (but not limited to) (1) the
    design of the robot, (2) the personal characteristics of the human user,
    and (3) the context and situation of use. For instance, a social context
    and purpose of the robot increase anthropomorphism: \textit{anthropomorphism
    as a multi-factor phenomenon}.

\item ~A person's tendency to anthropomorphize a robot
    evolves in a non-linear way over time and with growing experience that the
    person has with the robot. For instance, a more experienced user will
    perceive the robot as more familiar and predictable, and in turn
    anthropomorphizes the robot less: \textit{anthropomorphism as a dynamic
    phenomenon}.

\end{enumerate}

Further, we would like to investigate the link between the variances in anthropomorphism and the correlating cognitive states.

\subsection{Article overview}

%\begin{inparaenum}[\itshape 1)\upshape]

The remainder of this article is organized as follows.
According to the three perspectives that we take, 
we first (in Section
%\ref{sec:overview-studies}, we give a brief overview on the various studies that
%we conducted to investigate anthropomorphism, namely \item the Roomba study,
%\item the Forum study, \item the Ranger domino study. Next, Section
\ref{sec:anthropomorphism}) aim to establish a common ground in the
understanding of anthropomorphism. We review literature from several relevant domains and integrate the different explanations into an understanding of
anthropomorphism as a multi-factor phenomenon.
Based on this, Section~\ref{sec:our-ideas} introduces our initial model of the
\textit{Dynamics of Anthropomorphism}. This model describes the variances
and dynamics of people's tendency to anthropomorphize an artificial agent,
such as a robot. 
Then Section \ref{sec:cognition-neuroscience} presents a
cognitive and neuroscience perspective on anthropomorphism to consolidate
our considerations and the proposed model.
This is followed by a discussion
about the limitations of our approach (Section \ref{sec:discussion}) and the
existing and alternative ways to measure anthropomorphism in human-robot
interactions.
Finally, Section \ref{sec:conclusion} concludes this article
and offers starting points for further research on anthropomorphism and the
effect of the human-like design approach in robotics.

%\end{inparaenum}

>>>>

\documentclass{frontiersSCNS} % for Science articles

\usepackage[english]{babel}

%\setcitestyle{square}
\usepackage{url,lineno}
\linenumbers

\usepackage{datetime}
\usepackage{fmtcount}
\usepackage{etoolbox}

\usepackage{amsmath} 
\usepackage{mathptmx}

\usepackage{array}
\usepackage{graphicx}
%\usepackage{caption}
%\usepackage{subcaption}
\graphicspath{{figs/}}

\usepackage{color, soul}
\usepackage{url}
\usepackage{multirow}
\usepackage{array}
\usepackage{fixltx2e}
\usepackage{textcomp}

@@

@@

\def\journal{Cognitive Science}%%% write here for which journal %%%
\def\DOI{}
\def\articleType{Hypothesis and Theory}
\def\keyFont{\fontsize{8}{11}\helveticabold }
\def\firstAuthorLast{Fink {et~al.}} %use et al only if is more than 1 author
\def\Authors{Julia Fink\,$^{1*}$, S\'{e}verin Lemaignan$^{1}$, Claire Braboszcz$^{2}$ and Pierre Dillenbourg$^{1}$}
\def\Address{$^{1}$Computer-Human Interaction in Learning and Instruction (CHILI) \\ Ecole Polytechnique F\'{e}d\'{e}rale de Lausanne (EPFL) \\ CH-1015 Lausanne, Switzerland \\
\vspace{0.25cm}
$^{2}$Laboratory for Neurology and Imaging of Cognition (LabNIC) \\
Universit\'{e} de Gen\`{e}ve \\ CH-1211 Gen\`{e}ve, Switzerland 
}
\def\corrAuthor{Julia Fink \vspace{1mm}}
\def\corrAddress{EPFL - CHILI, RLC D1 740, Station 20, CH-1015 Lausanne, Switzerland}
\def\corrEmail{julia.fink@epfl.ch}

@@

%As a primary goal, the abstract should render the general significance and conceptual advance of the work clearly accessible to a broad readership. References should not be cited in the abstract.

\textit{Anthropomorphism} -- people's tendency to perceive human-like characteristics in non-human agents -- is a commonly observed phenomenon in Human-Robot Interaction (HRI). Though having been extensively studied, anthropomorphism is still not very well understood. Often, anthropomorphism is analyzed from the sole perspective of the human-like (anthropomorphic) design of a robot. 
%The question in focus is to what extent different types of anthropomorphic robot designs evoke affinity or uncanny feelings in the human user.
%In this regard, the \textit{Uncanny Valley} hypothesis suggests that an anthropomorphic design can evoke both positive and negative feelings in human users, determined by the degree of human-likeness. 
We believe that we need to go beyond this, in order to gain a better awareness about the phenomenon as a whole. To this end, we make three suggestions toward extending our understanding of anthropomorphism in HRI.
First, we suggest to comprehend anthropomorphism as a human-centered social phenomenon that arises in an interaction between a human and a robot in a specific context (partly determined by the purpose of the robot). Consequently, we propose that \textbf{multiple factors} need to be considered: the characteristics of the robot, of the human, and of the context. This holds implications for how anthropomorphism is studied and measured.
%Real physical interaction is required.
By applying a cognitive perspective, we further suggest that anthropomorphism is \textbf{multi-layered}, not being either \textit{there} or \textit{not there} but that there exist lower and higher levels of it, depending on the \textit{cognitive depth} that the attribution of a human-like characteristic to the robot implies, as well as on the specific mental model on which it is based.
Finally, we hypothesize that anthropomorphism is \textbf{dynamic}. In other words, it evolves over time, with growing real or imagined interaction experience and familiarization of the human user with the robot.
We formalize our ideas in a conceptual model that we call the \textit{Dynamics of Anthropomorphism}. Our hypotheses are based on results from our own empirical interaction studies, and informed on a review of related work in psychology, cognitive science and neuroscience as well as on previous research in HCI and HRI.
%As a consequence of our hypotheses, we discuss the methods and measurement tools to study anthropomorphism and propose going toward long-term interaction studies to verify our conceptual model.

\tiny
\keyFont{ \section{Keywords:} Anthropomorphism, Design, Dynamics, Human-Robot 
Interaction, Mental Models, Robot Acceptability, Social Robotics} 
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
%
%		INTRODUCTION
%
%
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Introduction}
\label{sec:introduction}

Anthropomorphism can be seen as a special type of
social response that people commonly show when interacting with 
robots. The phenomenon has been observed related to various types of technology, virtual agents, daily life objects, animals, and even simple geometric moving shapes. It seems that people have an almost natural tendency to \textit{``see human in the nonhuman''} -- and ascribe human-like characteristics to basically anything that acts or is believed to act with apparent independence \citep{epley_seeing_2007}. We focus here on anthropomorphism related to robots. In HRI it is commonly described that users tend to perceive and treat their robot as having some human-like characteristics, and hence attribute, for instance, emotional states, motivations, or intentions to it. Some Roomba (vacuum cleaning robot) owners even ascribe personality and gender to their robot and choose a name for it \citep{forlizzi_service_2006,forlizzi_how_2007,krumm_my_2007,sung_domestic_2010}.
To what extent people anthropomorphize a robot has so far mostly been studied from the sole perspective of the design of a robot. In short, the general assumption is that the more human-like a robot looks like and behaves, the more people will anthropomorphize it.
Not necessarily related to anthropomorphism but focusing on the uncanny aspect, Mori's \textit{Uncanny Valley} hypothesis suggests that the relation between the robot's degree of anthropomorphic form and people's feelings toward it is non-linear \citep{mori_uncanny_1970,mori_uncanny_2012}.
Somewhat similar to this non-linear relation between a robot's design and people's feeling toward it, we suspect that also people's tendency to anthropomorphize a robot is of dynamic nature. Furthermore, we hypothesize that the robot's design (its degree of anthropomorphic form) is not the only factor that plays a role in how much the robot is anthropomorphized by a user.

So far, the idea of multiple factors, as well as the dynamic nature of anthropomorphism have been mostly disregarded. There exists no research that investigated how anthropomorphism evolves over longer periods of time based on repeated physical interactions between the human and the robot. This is, however, a relevant aspect to investigate. Long-term interaction studies found that people's perception of a robot, their attitude toward it, how they use it, and how they experience it changes over time \citep{kanda_interactive_2004,karapanos_user_2009,sung_domestic_2010}. It is described that besides \textit{novelty effects} also the user's increasing \textit{familiarity} with the robot over time play a role and hence, we assume that this also impacts how much the robot will be anthropomorphized. We will shed light on this aspect by taking a (developmental / social) psychology and cognitive science perspective because these fields provide interesting explanations for people's tendency to anthropomorphize and we believe this can offer relevant extensions to our understanding of the phenomenon in HRI. Our general aim is to deepen our understanding of anthropomorphism, in order to advance research on this topic. We make three suggestions that can be addressed in future empirical research. Namely, we propose that anthropomorphism is based on multiple factors, that it is multi-layered, and that it is of dynamic nature.

%This is mostly theoretical material, gathered by reviewing diverse literature, reflecting upon it and integrating it. Anthropomorphism has first not been our main research topic, however, what led us to investigate the phenomenon in more detail, were surprising findings from our own HRI studies in which we observed anthropomorphism as an emerging phenomenon.
%We illustrate some of our empirical data and give concrete examples, so to stimulate the discussion about the current understanding of anthropomorphism, and guide the reader towards our ideas. One suggestion is, for instance, to measure anthropomorphism as being a characteristic of an interaction and not only of a user's perception of a robot.

%The proposed dynamics of anthropomorphism also implicate that we need to find new methods and tools to measure anthropomorphism, and we would like to take an initial step toward this by introducing a so-called \textit{anthropomorphism index}.

%
%Another reason may be that most of the research that examines the \textit{Uncanny Valley} is conducted in absence of real interaction but uses photos or videos of different robots instead. Furthermore, to measure anthropomorphism, mostly rating scales and questionnaires are applied. Based on our understanding of anthropomorphism we see a limitation in this, which we would like to discuss later.

%\footnote{To facilitate things, we use the terms \textit{`non-human 
%agent'}, \textit{`system'} and \textit{`robot'}, as well as 
%\textit{`artifact'} more or less interchangeably throughout this 
%article. Similarly, the terms \textit{`users'}, 
%\textit{`observers'}, \textit{`people'} and \textit{`humans'} are 
%used as synonyms.} 
%
%

%
%This idea is supported by a child-robot interaction study that we carried out with twenty-six 4-5 year old children.\footnote{A publication about this study is in work.} In a playful task, each two children were collaborating with our robotic box ``Ranger'' \citep{fink_which_2014} to assemble a domino. From time to time, the robot showed some unexpected behavior and we assessed to what extent children anthropomorphized the robot, \textit{e.g.} by ascribing intention or emotional states to robot.
%We found that children commonly attributed only some types of human-like characteristics, such as perceptual skills (which are not uniquely human), feeling, and moral standing. Also, children perceived the robot as a social entity and agreed it could be their friend. However, the majority of children did not ascribe intention to the robot ...

%So far in human-robot interaction (HRI), the study of anthropomorphism has mostly centered on the anthropomorphic design of a robot and its influence on people's perception of the robot. For instance, a prominent question is how more and less anthropomorphic robots (\textit{e.g.}, humanoid robot \textit{vs.} mechanic robot) are perceived in terms of human-likeness and to what extent the different designs evoke feelings of closeness / affinity or reluctance / revulsion. The so-called \textit{Uncanny Valley} hypothesis \citep{mori_uncanny_1970} states a non-linear relationship between these two variables, which has provoked a lot of research, which produced partly controversial findings, however.

%Not focusing on the uncanny aspect of a human-like robot design, a general question is whether people perceive and treat a human-like robot more readily as if it was a human. This is one of the major research topics in social robotics.

\subsection{What is Anthropomorphism?}
\label{sec:definitions}

Literature on anthropomorphism is diverse and there exist different and partly conflicting definitions (across and within disciplines). As a result of this, there are also different understandings of what exactly (which phenomenon) `anthropomorphism' describes.
%This may be due to the fact that anthropomorphism is studied in a variety of different domains that each integrate their own understanding and use the same terms to refer to different phenomena.
The problem is that despite the lack of a commonly accepted definition, the terms `anthropomorphism' and `anthropomorphic' are often used as if their meanings were clear and agreed upon \citep{persson_anthropomorphism_2000}. Especially in a multi-disciplinary field like HRI this can lead to a conflict. Thus, we have to be clear and specific about what we mean when we talk about anthropomorphism.

We propose to use the term `anthropomorphism' to refer to a specific \textit{human-centered social phenomenon} but to \textit{not} use the same term to refer to the amount of human-likeness in the \textit{design} of a robot (for latter one, we propose to use alternative formulations, such as `anthropomorphic design' or `human-like form'.) We believe that this differentiation is necessary and important. The point that we want to make is, that anthropomorphism arises in an interaction and as such, a robot does not \textit{``contain anthropomorphism''} or have a specific level of anthropomorphism in itself, but that it only gives rise to the process of being anthropomorphized by a given user. The ambiguity of terms has been previously discussed \citep{persson_anthropomorphism_2000,duffy_anthropomorphism_2002}.\footnote{Several attempts have been made to clarify what exactly anthropomorphism is. \cite{epley_when_2008} approached the issue from a psychological point of view give several examples of what anthropomorphism is \textit{not}. The reader may refer to the original work for more details.}

More specifically, we use the term `anthropomorphism' to describe a social phenomenon that arises in a real or imagined interaction between a human and a robot in a given context (see Figure~\ref{fig:multiple-factors}). The active player (the one who anthropomorphizes) is the human, and therefore, the phenomenon itself is human-centered. (Explanations of anthropomorphism can apply both a human-centered and an artifact-centered perspective, depending on the direction of causality which is studied: Is the human himself motivated to anthropomorphize the robot or is the robot, which leads the human to anthropomorphize it?) In our understanding, anthropomorphism includes both people's tendency to \textit{perceive / image} human-like characteristics in a robot, as well as the action of verbally \textit{ascribing / attributing} these characteristics to the robot and \textit{reacting / behaving / treating} the robot alike. Consequently, we believe that anthropomorphism can not only be assessed as a variable of a person's perception of the robot but also be observed in how he/she interacts with the robot (this holds implications for how we measure anthropomorphism, see Section~\ref{sec:anthropomorphism-index}). For instance, the user establishing eye-contact with a robot, using facial expressions when reacting to the robot, applying direct speech and polite formulations, or using social gestures toward the robot, may be counted as forms of anthropomorphism. In other words, we see anthropomorphism as a special type of \textit{human-like social engagement} with the robot (and to which the human commits).

\subsection{Overview}

The remainder of this article is organized as follows.

First, in Section~\ref{sec:material-methods} we give an overview of findings related to anthropomorphism from our own work. As several of our observations were rather surprising to us, we will use this to stimulate the reflection on the current understanding of anthropomorphism. Toward the end of this section, we propose a new concept towards measuring anthropomorphism, the so-called \textit{Anthropomorphism Index}, which renders a value that expresses a user's tendency to anthropomorphize a specific robot in a given context.

Section~\ref{sec:our-ideas} builds the main part of the article, which consists of the three suggestions that we make toward extending our understanding of anthropomorphism. We first, in Subsection~\ref{sec:multiple-factors}) outline how and why we think that anthropomorphism is based on multiple factors and not only (or mostly) on how much the robot resembles a human. Our argumentation is mostly driven by relevant theories from developmental and social psychology, as well as research on user experience and design.
As a second point, in Subsection~\ref{sec:multiple-layers}, we suggest that anthropomorphism is a multi-layered phenomenon, and for this we apply a cognitive perspective and further integrate relevant work from neuroscience.
Thirdly, Subsection~\ref{sec:dynamics} introduces the idea of how anthropomorphism evolves over time, when novelty effects wear off, and the user gets familiar with the robot and with using it. This is formalized in a conceptual model that we call the \textit{Dynamics of Anthropomorphism}.

(Section~\ref{sec:discussion}) discusses the role and effect of disruptive behaviors for anthropomorphism and in long-term HRI, and briefly presents ethical considerations.

Finally, in Section~\ref{sec:conclusions} we conclude the article. We  first summarize our main ideas to extend the understanding of anthropomorphism in HRI. Then, we present implications and limitations of our work.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
%
%		MATERIAL & METHODS
%
%
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\begin{methods}
\section{Material \& Methods}
\label{sec:material-methods}

%How did our own research and literature review lead us to the idea of developing this conceptual model?

In this section we give a brief overview of the studies that we carried out to
investigate anthropomorphism in human-robot interaction. These studies did not necessarily have anthropomorphism is the main topic, bu were motivated also by other research
questions, and we encourage the reader to refer to our other publications for more details about these studies.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
%
%		ROOMBA STUDY
%
%
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Roomba Study: changes in interaction and perception over time}
\label{sec:roomba-study}

We openly explored long-term interactions between human users and domestic
service robots in a 6-months lasting ethnographic study~\citep{fink_living_2013}. 
We gave commercially
available vacuum cleaning robots (iRobot's Roomba) to nine households.
There are two findings that are related to anthropomorphism.

\begin{figure}
\begin{center}
\includegraphics[width=10cm]{roomba-activities.pdf}% This is a *.jpg file
\end{center}
 \textbf{\refstepcounter{figure}\label{fig:roomba-activities} Figure \arabic{figure}.}{~~\small Over time there is a decrease of the social activities that households engaged in associated with the robot. Only 1 of the 9 households gave a name to their robot, and none of the households personalized their robot.}
\end{figure}

First, in terms of \textbf{anthropomorphic interactions} with the robot, we found that the effects were overall rather low and tended to fade out over time (first point is consistent with previous work by \cite{sung_housewives_2008}). Overall, we observed only few instances of people anthropomorphizing the robot or interacting with it in a human-social way. (Also, from the opposite point of view, the robot did not have a lasting social impact on the household, for instance, it did not lastingly change established cleaning roles.)
Some of the initially observed anthropomorphic interactions associated with the robot include \textit{watching} the robot for fun, giving a \textit{demo} of the robot to guests / visitors, \textit{personalizing} the robot, giving a \textit{name} to the robot, \textit{talking} to the robot using direct speech, and \textit{playing} with the robot. 
The majority of these anthropomorphic interactions with the robot did not last but participants stopped doing them after some time. Figure~\ref{fig:roomba-activities} shows how many of the 9 households engaged
in these activities during the 6 months when the study was conducted. We interpret that the nature of
human-robot interaction and how users experience the robot changes over time. We suspect that this is one of the reasons for why also the amount of
anthropomorphic / human-social interactions with the robot tends to decrease over
time. It is however possible that with some particular users, anthropomorphic interactions sustain.

\begin{figure}
\begin{center}
\includegraphics[width=14cm]{roomba-ratings.pdf}% This is a *.jpg file
\end{center}
 \textbf{\refstepcounter{figure}\label{fig:roomba-ratings} Figure \arabic{figure}.}{~~\small People's perception of the robot evolved over time, with stronger variations in the beginning. \textit{T1} = expectations, \textit{T2} = first usage, \textit{T3} = after 2 weeks, \textit{T4} = after 2 months, \textit{T5} = after 6 months.}
\end{figure}

Second, in terms of people's \textbf{perception of human-like characteristics and social traits of the robot}, we obtained similar findings. The robot was overall not perceived as having human-like characteristics, and attributions of intelligence, for instance, decreased over time.
At several time points during the 6-month lasting study, participants were asked to rate their perception and impression of the robot on 7-point Likert scales.
Concerning anthropomorphism, we included
\textit{perceived intelligence}, experienced \textit{fun} when using the
robot, and \textit{perceived impact} of the robot on the household as indicators for the perception of human-likeness in the robot. Figure~\ref{fig:roomba-ratings} shows how people's perception of the robot evolves over time.
Participants were generally not sure as how \textit{intelligent} they should rate the robot. Interestingly, at all five time points, male participants rated the robot as more intelligent (average 4.5) than females (average 3.4), however, this difference was not significant. Still, this suggests that gender, and probably other individual characteristics play a role. Also in terms of \textit{experienced fun} there was a non-significant but noteworthy gender difference. The rating for experienced fun remained quite stable for male participants, whereas it tended to decrease with time for female participants. The rating for \textit{perceived impact} reflects to what extent the robot is perceived to play a decisive role in the household. The average value remarkably increased after participants had used the robot the first time but then constantly decreased. This shows again, that people's perception of the robot changes over time, and that the initial moments using and interacting with a robot are special.
Overall, Figure~\ref{fig:roomba-ratings} shows that after the first usage of
the Roomba (T2), most of the perceived constructs slightly decrease (and hence
are less positively rated) as time goes by. The course is different for the
perceived \textit{ease of use} of the robot, as participants become more
experienced with the robot over time.

Summing it up, in accordance with previous work we also found relatively low rates of users
treating and perceiving their vacuum cleaning robot as a living (or human-like)
entity \citep{sung_housewives_2008}. We interpret that it might be the pure
functional purpose and non human-like form of the Roomba that prevented people
from anthropomorphizing it. Still, we observed some initial anthropomorphic
effects, both reflected in interaction with and perception of the robot.
However, most of the anthropomorphic effects did not last long. One fairly
stable human-like interaction with the robot was \textit{talking to the robot},
despite the fact that the robot was not able to recognize and respond to verbal
statements.
